Synthesizing Quotient Lenses

### 2018-03-13 papers/QuotientLenses/main.tex

To get a sense of how much the new expressiveness that combining QRE lenses with
synthesis saves in terms of programmer effort, we compared the amount of code
that would need to be written for each of the lens benchmarks. We use number of
nodes in the abstract syntax tree as a measure of code size, and calculated
three quantities for each lens synthesis benchmark task. We only compared the
sizes on benchmarks where the canonization functions are nontrivial.
%
\begin{itemize}
  \item[\QRESize{}] the number of AST nodes in the QRE
  specifications (including examples)\saz{CHECK THIS}\afm{no but i'll add it in}
  \item[\canonizeAndSpecSize{}]  the number of
  AST nodes in $W(q_1)$ summed with the number of AST nodes in $\canonize(q_1)$
  \item[\LensAndSpecSize{}] the number of AST nodes
  in $W(q_1)$ summed with the number of AST nodes in the synthesized QRE lens.
\end{itemize}

@@

Figure~\ref{fig:asts} shows the number of benchmarks that can be expressed using
at most a given number of AST nodes. For a given benchmark, we found \QRESize{}
to be about half as large as \canonizeAndSpecSize, and about a third as large
as \LensAndSpecSize{}.\BCP{These are really the critical conclusions, and they
  are hard to see, the way the graph is drawn. I wonder whether we could instead
  present with the individual examples on the X axis (sorted by number of AST
  nodes), with a bar above each one containing three regions, colored
  differently, for LS, CS-LS, and QC-CS...}\afm{will do} This suggests that
introducing QREs saves programmers significant effort compared to both Optician
and basic Boomerang.

>>>>

To get a sense of how much the new expressiveness that combining QRE lenses with
synthesis saves in terms of programmer effort, we compared the amount of code
that would need to be written for each of the lens benchmarks. We use number of
nodes in the abstract syntax tree as a measure of code size, and calculated
three quantities for each lens synthesis benchmark task. We only compared the
sizes on benchmarks where the canonization functions are nontrivial.
%
\begin{itemize}
  \item[\QRESize{}] the number of AST nodes in the QRE
  specifications (including examples).
  \item[\canonizeAndSpecSize{}]  the number of
  AST nodes in $W(q_1)$ summed with the number of AST nodes in $\canonize(q_1)$
  and the number of examples.
  \item[\LensAndSpecSize{}] the number of AST nodes
  in $W(q_1)$ summed with the number of AST nodes in the synthesized QRE lens.
\end{itemize}

@@

Figure~\ref{fig:asts} shows the number of benchmarks that can be expressed using
at most a given number of AST nodes. On average, \canonizeAndSpecSize{} used
44\% more AST nodes than \QRESize{}, using an average of 237 more AST nodes. On
average, \LensAndSpecSize{} used 193\% more AST nodes than \QRESize{}, using an
average of 1267 more AST nodes. This suggests that introducing QREs saves
programmers significant effort compared to both pre-QRE Optician and basic
Boomerang.
